---
title       : Loan Repayment App
subtitle    : Course Project - Developing Data Products
author      : 
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme: solarized_light
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---   

## Planning to buy something expensive?

- Thinking of buying something expensive like a car or a house?  
- But do not have enough money to pay upfront  
- You probably will look at getting a loan  
- [Loan Repayment Application](https://bmat.shinyapps.io/LoanRepayment/) provides a simple way to find some of the information that you need to make you decision 

--- .class #id 

## A simple and easy way to work out details of the loan

- [Loan Repayment Application](https://bmat.shinyapps.io/LoanRepayment/) is a very simple tool to work out the details of loan that you might want to borow  
- It allows you to play around with the key variables in borrowing
- And provide you the information on the contribution of you payment towards interest and the loan principal amount
- <span class = 'red'>Note: </span> Do consult your lender of other options they may have before finalising you decision


--- &twocol w1:50% w2:50%

## Check out the charts ...  

- Charts showing proportion towards interest and principal for 30 and 20 year loans
- The area in red shows the proportion that you pay towards interest; Longer the loan term, higher the proportion that goes to interest.

*** =left

```{r echo=FALSE, eval=TRUE, fig.width=6, fig.height=6}
pv<-150000
rate<-8/1200
years<-30
per <- years * 12
pmt <- (rate + (rate / (((1+rate)^per) - 1) )) * pv
bdw <- data.frame(year=numeric(per), period=numeric(per), 
                  repayment=numeric(per), 
                  toPrincipal=numeric(per), 
                  percentToPrincipal=numeric(per),
                  remainingPrincipal=numeric(per))
bdw$repayment <- pmt
remPrincipal <- pv
for(i in 1:per) {
    bdw$year[i] <- format(ceiling(i/12),nsmall=0)
    bdw$period[i] <- format(i, nsmall=0)
    bdw$toPrincipal[i] <- P <- pmt-(remPrincipal * rate)
    bdw$percentToPrincipal[i] <- (P / pmt) * 100
    remPrincipal <- remPrincipal - P
    bdw$remainingPrincipal[i] <- remPrincipal
    }
#adjustment for last payment
bdw$repayment[per]<-pmt+bdw$remainingPrincipal[per]
bdw$remainingPrincipal[i] <- 0

totalInterest <- sum(bdw$repayment) - sum(bdw$toPrincipal)
totalPayment <- sum(bdw$repayment)
totalToPrincipal <- sum(bdw$toPrincipal)

with(bdw, plot(period, toPrincipal, type="n", 
               ylab="Contribution of repayment ($ or currency entered)",
               xlab="Period (months)",
               main="Repayment contribution to principal and interest",
               ylim=c(0, repayment[1])))

polygon(c(1, bdw$period, dim(bdw)[1]), c(0, bdw$toPrincipal, 0),
        col="skyblue")
polygon(c(bdw$period[1], bdw$period, bdw$period[length(bdw$period)]), 
        c(min(bdw$repayment), bdw$toPrincipal, min(bdw$repayment)),
        col="red")
legend("bottomright", pch=15, col = c("red", "skyblue"), 
       legend = c("Pay towards interest", "Pay towards principal"))

text((0.194)*as.numeric(bdw$period[length(bdw$period)]), (0.955)*bdw$repayment, 
     paste0("Interest - $", format((totalInterest), nsmall=0)))

text((0.861)*as.numeric(bdw$period[length(bdw$period)]), 
     (0.177)*bdw$repayment, paste0("Principal - $", 
                                  format(totalToPrincipal, nsmall=0)))

text( (0.5)*as.numeric(bdw$period[length(bdw$period)]), 
      (0.5)*bdw$repayment, paste0("Total payment - $", format(totalPayment, nsmall=0)))

```


*** =right

```{r echo=FALSE, eval=TRUE, fig.width=6, fig.height=6}
pv<-150000
rate<-8/1200
years<-20
per <- years * 12
pmt <- (rate + (rate / (((1+rate)^per) - 1) )) * pv
bdw <- data.frame(year=numeric(per), period=numeric(per), 
                 repayment=numeric(per), 
                 toPrincipal=numeric(per), 
                 percentToPrincipal=numeric(per),
                 remainingPrincipal=numeric(per))
bdw$repayment <- pmt
remPrincipal <- pv
for(i in 1:per) {
    bdw$year[i] <- format(ceiling(i/12),nsmall=0)
    bdw$period[i] <- format(i, nsmall=0)
    bdw$toPrincipal[i] <- P <- pmt-(remPrincipal * rate)
    bdw$percentToPrincipal[i] <- (P / pmt) * 100
    remPrincipal <- round(remPrincipal - P, 2)
    bdw$remainingPrincipal[i] <- remPrincipal
    }
#adjustment for last payment
bdw$repayment[per]<-pmt+bdw$remainingPrincipal[per]
bdw$remainingPrincipal[i] <- 0

totalInterest <- sum(bdw$repayment) - sum(bdw$toPrincipal)
totalPayment <- sum(bdw$repayment)
totalToPrincipal <- sum(bdw$toPrincipal)

with(bdw, plot(period, toPrincipal, type="n", 
              ylab="Contribution of repayment ($ or currency entered)",
              xlab="Period (months)",
              main="Repayment contribution to principal and interest",
              ylim=c(0, repayment[1])))

polygon(c(1, bdw$period, dim(bdw)[1]), c(0, bdw$toPrincipal, 0),
        col="skyblue")
polygon(c(bdw$period[1], bdw$period, bdw$period[length(bdw$period)]), 
        c(min(bdw$repayment), bdw$toPrincipal, min(bdw$repayment)),
        col="red")
legend("bottomright", pch=15, col = c("red", "skyblue"), 
       legend = c("Pay towards interest", "Pay towards principal"))

text((0.194)*as.numeric(bdw$period[length(bdw$period)]), (0.955)*bdw$repayment, 
     paste0("Interest - $", format((totalInterest), nsmall=0)))

text((0.861)*as.numeric(bdw$period[length(bdw$period)]), 
     (0.177)*bdw$repayment, paste0("Principal - $", 
                                  format(totalToPrincipal, nsmall=0)))

text( (0.5)*as.numeric(bdw$period[length(bdw$period)]), 
      (0.5)*bdw$repayment, paste0("Total payment - $", format(totalPayment, nsmall=0)))
paymentBreakDown <- bdw
```


---

## ... and the table with the breakdown   

   
```{r echo=TRUE, results='asis'}
library(xtable)
print(xtable(paymentBreakDown[1:3,]), type="html") #display first few lines
```
  
```{r echo=TRUE, results='asis'}
print(xtable(paymentBreakDown[238:240, ]), type="html") #display last few lines
```




